---
title: "ch 7 R review"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

load("7.R.RData")

```

Q1
Load the data from the file 7.R.RData, and plot it using plot(x,y). What is the slope coefficient in a linear regression of y on x (to within 10%)?


```{r}
plot(x, y)

all <- as.data.frame(cbind(x, y))

lmod <- lm(y ~ x, data = all)

summary(lmod)
```

Q2
For the model y ~ 1+x+x^2, what is the coefficient of x (to within 10%)?

```{r}
lmod2 <- lm(y ~ 1 + x + poly(x, 2, raw = T), data = all) #if raw = T is not set, then orthogonal polynomial will be given, which is not what you want (not sure why) -- in the first R video (min 3:20) it was like in regression with uncorrelated predictors

summary(lmod2)

#alternative answer
lmod3 <- lm(y ~ 1 + x + I(x^2))

summary(lmod3)
```

7.Q.1
Suppose we want to fit a generalized additive model (with a continuous response) for y against X1 and X2. Suppose that we are using a cubic spline with four knots for each variable (so our model can be expressed as a linear regression after the right basis expansion).

Suppose that we fit our model by the following three steps:

1) First fit our cubic spline model for y against X1, obtaining the fit f^1(x) and residuals ri=yi−f^1(Xi,1).

2) Then, fit a cubic spline model for r against X2 to obtain f^2(x).

3) Finally construct fitted values y^i=f^1(Xi,1)+f^2(Xi,2).

Will we get the same fitted values as we would if we fit the additive model for y against X1 and X2 jointly?

Answer: Not necessarily, even if x1 and x2 are uncorrelated. If X1 and X2 are uncorrelated, and we are only fitting a linear regression (only linear terms for X1 X2 ) then this method would work. However, even if X1 and X2 are uncorrelated, the nonlinear basis functions might be.

```{r}
set.seed(42)

x1 <- rnorm(1000)
x2 <- abs(x1)
x3 <- x1 +2
y <- rnorm(1000) + rnorm(1000)

cor(x1, x2) #uncorrelated
cor(x1, x3) #perfect correlation

dat <- as.data.frame(cbind(x1, x2, y))

#1st step specified in the question
gammod <- lm(y ~ bs(x1, df = 5), data = dat) #as per my understanding, df = 5 because the question specificied 4 knots and i think the formula for the knot is df-degree (1 if intercept). so to get 4 knots, need 5 df.
summary(gammod)
dat$pred <- predict(gammod, newdata = dat)
dat$resid <- dat$y - dat$pred

dat

#2nd step
gammod2 <- lm(resid ~ bs(x2, df = 5), data = dat)

dat$pred2 <- predict(gammod2, newdata = dat)
dat$resid2 <- dat$resid - dat$pred2

dat

#3rd step
dat$finalfit <- dat$pred + dat$pred2
dat

#fitting the model in 1 go (jointly)
jointmod <- gam(y ~ s(x1, df = 5) + s(x2, df = 5), data = dat)
dat$pred3 <- predict(jointmod, newdata = dat)
dat

```

If X1 and X2 are uncorrelated, and we are only fitting a linear regression (only linear terms for X1 X2 ) then this method would work. However, even if X1 and X2 are uncorrelated, the nonlinear basis functions might be.
below is the example for linear regression. pred3 and finalfit are very similar (x1 and x2 have very small correlation)
```{r}
set.seed(42)

x1 <- rnorm(1000)
x2 <- abs(x1)
x3 <- x1 +2
y <- rnorm(1000) + rnorm(1000)

cor(x1, x2) #uncorrelated
cor(x1, x3) #perfect correlation

dat <- as.data.frame(cbind(x1, x2, y))

#1st step specified in the question
gammod <- lm(y ~ x1, data = dat) #as per my understanding, df = 5 because the question specificied 4 knots and i think the formula for the knot is df-degree (1 if intercept). so to get 4 knots, need 5 df.
summary(gammod)
dat$pred <- predict(gammod, newdata = dat)
dat$resid <- dat$y - dat$pred

dat

#2nd step
gammod2 <- lm(resid ~ x2, data = dat)

dat$pred2 <- predict(gammod2, newdata = dat)
dat$resid2 <- dat$resid - dat$pred2

dat

#3rd step
dat$finalfit <- dat$pred + dat$pred2
dat

#fitting the model in 1 go (jointly)
jointmod <- lm(y ~ x1 + x2, data = dat)
dat$pred3 <- predict(jointmod, newdata = dat)
dat

```

